Comparing probabilistic predictive models applied to football

### 2015-06-30 template.tex
secao 3 revisada

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

@@

How can we fairly distinguish better-informed states of knowledge from more poorly informed states?
A way of doing this is to use proper scoring rules for evaluating states of uncertain knowledge, where
the score may be interpreted as a numerical measure of how inaccurate was the prevision.

%More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome $X$ and its respective prevision value, $P(X)$.
%It is maximum for each $x$ when $x=P(X)$, and is nonincreasing both as $x$ departs from $P(X)$ (for each $P(X)$) and as $P(X)$ departs from $x$ (for each $x$).
%Thus, the best score is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be. Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S[x,P(X)]=0$ for each $x$.

%Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P(X)$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
%Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

%Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P(X)$ that minimizes $E_{P_0}[S(X,P(X))]$ is $P_0$.
%In other words, if one receives a reward $S(x,P(x))$ when $x$ happens and if $S$ is a proper scoring rule, then the highest expected reward is obtained by reporting one's  true uncertainty about $X$.
%The use of a proper scoring rule thus encourages the forecaster to be honest to maximize the expected reward.
%Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events.
%Therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome of $X$ and its respective prevision value, $P_X$, i.e. $P_X$ is the probability assigned to each possible outcome of $X$.
It is minimum for each $x$ when $P_X$ is such that $P_X(X=x)=1$, and is nondecreasing both as $x$ departs from $P_X$ (for each $P_X$) and as $P_X$ departs from $x$ (for each $x$).
Thus, the best score (the smaller value of the score) is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good predictions.}.
%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P_X$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P_X$ that minimizes $E_{P_0}[S(X,P_X)]$ is $P_0$.
In other words, if one receives a penalty $S(x,P_X)$ when $x$ happens and if $S$ is a proper scoring rule, then the lowest expected penalty is obtained by reporting one's  true uncertainty about $X$.
The use of a proper scoring rule thus encourages the forecaster to be honest to minimize the expected penalty.
Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

@@

Let $P_X$ be a probability assignment for $X$.
The Brier score is given by
$$S(X;P_X)= \sum_{i}\mathbb{I}(X=x_i)(1- P_X(x_i))^2+\sum_{i}\mathbb{I}(X\neq x_i)P_X^2(x_i),$$
where $\mathbb{I}$ is the indicator function.

@@

\end{tikzpicture}
\caption{Bi-dimensional simplex (green surface): area of possible predictions}\label{fig:simplex}
\end{center}
\end{figure}

We interpret the Brier score in the case where one of three mutually exclusive outcomes happen, the case of a football match.
The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all points of the simplex.
This new triangle is used to graphically display the predictions as internal points because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

The Brier score for the prediction $P_X=(0.25,0.35,0.40)$, assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
If the prediction was $P_X=(0,0,1)$, the score would be zero, the minimum for this rule.

It is useful to consider the score of what we will call {\it trivial prediction}: $(1/3,1/3,1/3)$.
This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

The logarithmic  score is given by
$$S(X;P_X)=- \sum_{i}\mathbb{I}(X=x_i)\ln(P_X(x_i)),$$

@@

The logarithmic score for the prediction $P_X=(0.25,0.35,0.40)$, when the home team loses is therefore 
$-\ln(0.4)\approx 0.91$. 
If the prediction was $P_X=(0,0,1)$, the score would be zero, once again the minimum of this rule.
For the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

The spherical score is given by
$$S(X;P_X)=- \frac{1}{\sqrt{\sum_i P_X^2(x_i)}}\sum_{i}\mathbb{I}(X=x_i)P_X(x_i),$$

@@

The spherical score for the prediction $P_X=(0.25,0.35,0.40)$, assuming the home team loses, is given by
$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
If the prediction was $P_X=(0,0,1)$, the score would be $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
More precisely, the proportion of errors of a given prediction $P_X$ is defined by
$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x} P_X(x_{j})\right),$$
where $X_j$ is the outcome of the $j$-th match and $P_X(x_{j})$ is the probability attributed to result $x\in\mathcal{X}=\{1,2,3\}$ for the $j$-th match.

@@

%De Finetti used what is called the Brier score, which is just the (squared) Euclidean distance — i.e., the sum of the squared differences between credences and (the 0/1 numerical correlates of) truth values.

>>>>

%  LaTeX support: latex@mdpi.com\textbf{}
%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

@@

	How can we fairly rank the various predictive models?
	A way of doing this is by using proper scoring rules, where
	the score may be interpreted as a numerical measure of how inaccurate was a given probabilistic prediction.

	Formally, let $X$ be a random variable indicating
	the result of a given match. Hence, $X$ assumes a value in $\{1,2,3\}$. Moreover, let $P$ denote one's probabilistic prediction
	about $X$, i.e., $P$ is such that $\sum_{x=1}^3P(X=x)=1$ and
    $P(X=x)\geq0$ for all $x$ in $\{1,2,3\}$.
	A scoring rule is a function 
	that assigns a real number (score) $S(x,P)$ to each combination $(x,P)$, where $x$ denotes each possible outcome of the match.
	For a given $x$ in $\{1,2,3\}$, the score  $S(x,P)$ is minimum when $P$ is degenerate distribution on $x$, i.e., when $P(X=x)=1$.  
	Thus, the best score (i.e., the score with the smallest value) is achieved when the probabilistic prediction for the outcome of the game is a degenerate  distribution
  on the observed outcome of that match, i.e., a perfect prediction.\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good forecasts.}
	%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

	Although many functions satisfy the criteria for a scoring rule specified above, 
	not all scoring rules encourage honesty and accuracy in assessing numerical values for their predictions.
	Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}, which we describe in the sequence.

	Say a model's probabilistic prediction for $X$ is $P^*.$ A proper scoring rule is a scoring rule $S$ such that the probabilistic prediction $P$ that minimizes 
	$$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*(X=x)$$
	 is $P^*$  itself.
	In other words, if one imagines 
	a penalty $S(x,P)$ is received 
	when $x$ happens then, 
	if $S$ is a proper scoring rule, the lowest expected penalty is obtained by reporting the model's  uncertainty about $X$, $P^*$.
	The use of a proper scoring rule thus encourages the forecaster to use $P^*$
	as his probabilistic prediction for the outcome of the game, $X$, rather than some other quantity. 
	%Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

	Let $P$ be a probabilistic prediction  for $X$.
	The Brier score for a given outcome $x\in\{1,2,3\}$ is given by
	$$S(x,P)= \sum_{i=1}^3\mathbb{I}(x=i)(1- P(X=i))^2+\sum_{i=1}^3\mathbb{I}(x\neq i)P^2(X=i),$$
	where $\mathbb{I}$ is the indicator function.

@@

			\end{tikzpicture}
			\caption{Bi-dimensional simplex (green surface): area of possible forecasts}\label{fig:simplex}
		\end{center}
	\end{figure}

	We interpret the Brier score in the case where one of three mutually exclusive outcomes happen, the case of a football match.
	The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
	The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
	Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all points of the simplex.
	This new triangle is used to graphically display the forecast as a internal point because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

	The Brier score for the probabilistic prediction
	$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
	assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
	On the other hand, the prediction $P(X=3)=1$ achieves score  zero, the minimum for this rule.

	It is useful to consider the score of what we will call {\it trivial prediction}:
	$$P(X=1)=1/3, P(X=2)=1/3, P(X=3)=1/3.$$
	This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

	The logarithmic  score is given by
	$$S(x,P)=- \sum_{i=1}^3\mathbb{I}(x=i)\ln(P(X=i)),$$

@@

	The logarithmic score for the prediction 
		$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
		 when the home team loses is therefore 
	$-\ln(0.4)\approx 0.91$. 
	On the other hand, the prediction $P(X=3)=1$ achieves score zero, once again the minimum of this rule.
	Moreover, for the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

	The spherical score is given by
	$$S(x,P)=- \frac{1}{\sqrt{\sum_{i=1}^3 P^2(X=i)}}\sum_{i=1}^3\mathbb{I}(x=x_i)P(X=i),$$

@@

	The spherical score for the prediction 
		$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
		 assuming the home team loses, is given by
	$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
	On the other hand, the prediction $P(X=3)=1$ achieves score  $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

	The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
	More precisely, the proportion of errors of a given probabilistic prediction $P$ is defined by
	$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x \in \{1,2,3\}} P(X_j=x)\right),$$
	where $X_j$ is the outcome of the $j$-th match.

@@

	%De Finetti used what is called the Brier score, which is just the (squared) Euclidean distance ? i.e., the sum of the squared differences between credences and (the 0/1 numerical correlates of) truth values.
